Case study: a spatial conservation plan for the Hunter region, NSW, Australia
=============================================================================

To demonstrate how to incorporate the value of information in a systematic spatial conservation plan, we now turn to a case study on priortizing the Hunter region for the conservation of threatened plants and animals. For the case study we make the simplifying assumptions that the entire region is an original position where no area is protected but the entire region is available for protection in a conservation plan. While this is entirely unrealistic, it would be unneccessary to complicate the demonstration with a more realistic scenario as the complication would only serve to distract the reader from the key components of the method we outline here.

Study area
----------

The Hunter is a biodiverse region of north-eastern New South Wales, Australia. The region is home to many threatened species of plants and animals. There are multiple threats to biodiversity in the region. The Hunter is under is under active development and the area's land users utilize it's resources for mining, agricultrue, transport, urban infrastructure and conservation. For the the analyses we present here, we consider the Hunter region to include the local government areas of Cessnock, Dungog, Gloucester, Gosford, Greater Taree, Great Lakes, Lake Macquarie, Maitland, Musselbrook, Newcastle, Port Macquarie-Hastings, Port Stephens, Singleton, Upper Hunter and Wyong, an are of `r fmt(cellStats(!is.na(hunter_alt), sum))` km$^2$.

Study species
-------------

The Hunter is home to many species of national conservation significance. Here we consider six species: two birds, two mammals and two plants. For the following analyses we build conservation plans that aim to maximise the average relative carrying capacity of these six species across the hunter region. Table 1 outlines these six species and an estimate of their maximum carrying capacity (see supplement for more details) as well as their conservation status according to the _NSW Threatened Species Conservation Act 1995_ (TSC), _Commonwealth Environment Protection and Biodiversity Conservation Act 1999_ (EPBC) and the IUCN Red list (IUCN).

| Common Name          | Scientific Name         |   $\bar{K}^\mathrm{max}$ | TSC | EPBC | IUCN |
|----------------------|-------------------------|-------------------------:|-----|------|------|
| Powerful Owl         | _Ninox strenua_         |                      0.1 | V   | -    | LC   |
| Spotted-tailed Quoll | _Dasyurus maculatus_    |                      0.2 | V   | E    | NT   |
| Regent Honeyeater    | _Anthochaera phrygia_   |                      2.0 | CE  | CE   | CE   |
| Squirrel Glider      | _Petaurus norfolcensis_ |                    150.0 | V   | -    | LC   |
| Bynoe's Wattle       | _Acacia bynoeana_       |                    250.0 | E   | V    | -    |
| Charmhaven Apple     | _Angophora inopina_     |                  18000.0 | V   | V    | -    |

Table: Species used in the conservation plan for the Hunter region. LC = Least Concern;  NT = Near Threatened; V = Vulnerable; E = Endangered; CE = Critically Endangered. $\bar{K}^\mathrm{max}$ = Estimated maximum carrying capacity (no individuals per square kilometer).

<!---
## Approximate maximum carrying capacity of the Hunter region

Powerful Owl home range is [1e4 to 5e4 hectares](http://www.publish.csiro.au/paper/MU06055.htm). Maximum Carrying capacity is probably something > .001 birds per hectare.

Tiger Quoll home range is [5e3 to 4e4 hectares](http://onlinelibrary.wiley.com/doi/10.1017/S0952836903004631/abstract). Maximum Carrying capacity is probably something > .002 quolls per hectare.

Squirrel Glider maximum carrying capacity is [1.5 gliders per hectare](http://www.alburyconservationco.org.au/wp-content/uploads/2012/10/Population-Viability-Analysis-for-Squirrel-Gliders-in-Thurgoona-NSW_20091.pdf).

Regent honeyeater: according to [this](http://www.publish.csiro.au/paper/PC990224.htm)
0.12 (+-.72) birds per hectare (not normally distributed though). So a rough guess may be 2 birds per hectare.

Angophora inopina: to [this](http://www.publish.csiro.au/paper/BT03058.htm) 1100 trees where found in 6.2 hectare. So approximately 180 plants per hectare.

Acacia bynoeana: [1600 plants](https://www.environment.gov.au/cgi-bin/sprat/public/publicspecies.pl?taxon_id=8575) / [650ha](http://www.environment.nsw.gov.au/resources/parks/PoMLakeMacquarieSCAPulbahIslandNRMoonIslandNR.pdf)So approximately plants per 2.5 plants per hectare.
--->

Input data for the conservation plan
------------------------------------

### Predictors of species distributions

We summarise the environment of the Hunter region with six data layers: annual mean solar radiation, annual mean temperature, annual precipitation, precipitation seasonality (coefficient of variation), inherent soil fertility, and topographic wetness index (Figure 3). Each layer is a `r nrow(predictors)` by `r ncol(predictors)` grid of 1 km$^2$ cells. We chose this set of variables as they are publically available (see suplement for sources), are biologically plausible drivers of the distribution of many taxa, have previously been shown to predict the distributions of the study species in the region [**cite**] and are relatively uncorrelated with one another (maximum pearson correlaton coefficient = `r  fmt(max(abs(sort(layerStats(predictors, "pearson", na.rm = TRUE)[[1]], TRUE)[-1:-nlayers(predictors)])), 2)`).

### Species occurence data

We obtained 30 random occurrence records within the boundaries of the Hunter region (as defined above) and within the date range, 1996 to 2016, for each of the six study species from the Atlas of Living Australia database [**cite**]. We chose this relatively low sample-size random subset to ensure that subsequent modelling would have an initial level of uncertainty large enough for us to demonstrate how to calculate the value of information.

```{r box2, child = "box2.rmd"}
```

Modelling species distributions using Maxent
--------------------------------------------

We used the modeling software MaxEnt [**cite**] to create species distribution maps on which to base the conservation plan. MaxEnt can be used to describe the potential distribution of species with occurrence records alone. The algorithm does this by comparing the distribution of occurrence records in covariate space to the distribution of covariate space as a whole, also known as the background (see below). We fit MaxEnt models to each species 

### Background covariate data

```{r box3, child = "box3.rmd"}
```

| Expected value | Average proportion of \\ carrying capacity remaining |
|----------------|------------------------------------------------------|
| EVWOI          | `r fmt(evwoi * 100)`                                 |
| EVWPI          | `r fmt(evwpi * 100)`                                 |
| EVPI           | `r fmt((evwpi - evwoi) * 100)`                       |

Table: Something
